---
title: "Practical Machine Learning"
author: "Romain Jacotin"
date: "26 avril 2015"
output: html_document
---

# Course project

## Make tidy data

```{r}
# Reading data
training <- read.csv("pml-training.csv")
testing  <- read.csv("pml-testing.csv")

# Cleaning data

# Removing all features containing 'NA' values int training and/or testing data set
n            <- dim(training)[2]
featureIndex <- rep(TRUE,n)
for (i in 1:n ) {
    featureIndex[i] <- (sum(is.na( testing[,i] ))>0) || (sum(is.na( training[,i] ))>0)
}

# Removing not useful features:
# "X"
featureIndex[1] <- TRUE
# "user_name"
featureIndex[2] <- TRUE
# "raw_timestamp_part_1"
featureIndex[3] <- TRUE
# "raw_timestamp_part_2"
featureIndex[4] <- TRUE
# "cvtd_timestamp"
featureIndex[5] <- TRUE
# "new_window"
featureIndex[6] <- TRUE
# "num_window"
featureIndex[7] <- TRUE

# Cut each training and testing data sets in two : the dirty features part and the tidy features part
testingNA  <- testing[ , featureIndex ]
trainingNA <- training[ , featureIndex ]

testingTidy  <- testing[ , !featureIndex ]
trainingTidy <- training[ , !featureIndex ]
```

## Partition original data into a Training and Testing data set

```{r}
library(caret)
inTrain <- createDataPartition( trainingTidy$classe, p=0.7, list=FALSE )
train   <- trainingTidy[ inTrain, ]
test    <- trainingTidy[ -inTrain, ]

# Five classes repartition (A/B/C/D/E) in the newly created train (70%) and test set (30%)
table(train$class)
table(test$class)
```

## Fit a Random Forest model

```{r}
# modFit  <- train( classe~., data=train )

```


## Use model to predict


predTrain <- predict( modFit, train )
predTest  <- predict( modFit, test )


## Evaluate model


confusionMatrix( predTrain, train$classe )
confusionMatrix( predTest, test$classe )


## Use model to predict testing for submission


predSubmit <- predict( modFit, testingTidy )


## Create independant files for submission


pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}

submitData <- as.character(predSubmit)
pml_write_files(submitData)


